What words ought to exist? \\ {\normalsize Coining with coinduction}

### 2011-03-28 trunk/wishlist/paper.tex
Improved for b&w printing

% \usepackage{eclbkbox}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{amscd}
% \usepackage{xy}
\usepackage{graphicx}
% \usepackage{fancyhdr}
% \usepackage{color}
% \usepackage[dark,all,bottom,landscape,timestamp]{draftcopy}
% \usepackage{everypage}

@@

\date{3 March 2011}

@@

\begin{abstract}
what do I put here
\end{abstract}

\vspace{1em}
{\noindent \small {\bf Keywords}:
 cryptolexicography, n-Markov models, coinduction
}

\section*{Introduction}
During a recent high-stakes game of Scrabble-brand Crossword
Puzzle\footnote{Scrabble is a registered trademark of Hasbro
Inc./Milton Bradley, and Mattel/JW Spear \& Sons plc.} I had what
could only be described as a killer bingo word (all 7 tiles) that,
after careful study, I determined could not be placed anywhere on the
board. Later in that same game, I had another sequence of letters that
just totally seemed like it should be able to make some long-ass
words, like for example ``oilsoap'' which turns out is not a legal
Scrabble word.\!\footnote{There are actually no 7-letter words that
can be made from these letters. Don't even bother. Even if playing off
an existing letter on the board, the best we can do are the non-bingos
``topsoil,'' ``topsail,'' or ``poloist'' with an available {\it t}.}
This naturally made me frustrated and I wanted to do something about
it. Why can't ``oilsoap'' be a word? Or ``loopsia''? Words are
introduced into the lexicon all the time. My first reaction of course
was to make an online version of Scrabble where all words are legal.
This is available at {\tt http://snoot.org/toys/scrallbe}, and is
pretty boring, I gotta be honest.
% TODO: Screenshot of Scrallbe

@@

\begin{figure}
\includegraphics[width=\linewidth]{wishlist-cdf}
\caption{Distribution of word frequency. Approximately 25,000
  different words (y axis) were issued 55 times or fewer (x axis).}
\label{fig:gamedistribution}
\end{figure}

@@

The next step is to build a Markov chain for a list of real words and
see what results. I built one for the SOWPODS word list, and these
are the most likely words, with real words filtered out:

% XXX improve layout
\vspace{0.2in}
\begin{tabular}{rl@{\quad\quad}rl}
0.049901545  & s     &   0.001768358  & y   \\
0.017526947  & d     &   0.001761301  & p   \\
0.009521421  & g     &   0.001653041  & a   \\
0.005506083  & c     &   0.001652864  & n   \\
0.004376836  & r     &   0.001563810  & ps  \\
0.004218223  & t     &   0.001383503  & ms  \\
0.004088377  & e     &   0.001371545  & ts  \\
0.003560976  & m     &   0.001327023  & ds  \\
0.003248559  & ss    &   0.001188715  & hy  \\
0.002001529  & rs    &   0.001166457  & k   \\
0.001958674  & h     &   0.001160619  & ng  \\
0.001873761  & l     &   0.001142785  & ly  \\
\end{tabular}
\vspace{0.2in}

@@

There is a standard solution to this problem, which is to generalize
the Markov chain to keep more than one letter of history. So instead
of just tallying how often {\it g} follows {\it n}, we count how often
{\it g} follows {\it in} (and any other pair of letters). This makes
the table pretty large, but I went through some trouble to put a
graphic in Figure~\ref{fig:sowpodstrigrams} because it looks pretty
cool.

% laplace smoothing?
% TODO: the empty string?

TODO: most likely words in italian, too.

most people remember induction, perhaps

coinduction: never gonna give you up. once you pop you can't stop.

% word list from Italian Scrabble, which I bet is called
% {Scrabbizzimo!}

@@

% \section{Ockham's Regexp?}

@@

On occasion I have been accused of ``overthinking''

Rob: etsy, nuuog
Chris: tapping on the bisphenol-A, nurm
David: wafflucinations
Lea: hnfff sound
Reed: pansepticon
Jessica: gruntle (like not disgruntled)

In order to not reprint everyone's bullshit---but not introduce bias by selectively removing data---I discarded random subsets of the data until it did not contain bullshit any more.

% TODO: backformation, stripping dis- and anti- and un- from words
% then checking to see if they're still words?

@@

\section{Recommendations}

sweeeeeeeeeeeeeeeeeeet with 19 {\it e}s, which means ``Really sweet.''

``rane'' sounds too much like ``rain'', but ``sare'' has a unique pronunciation and many people seem to think it's already a word. ``cho'' is similarly easy to pronounce and spell. I propose that it be defined as ``A kind of cheese,'' so that we can really nail the new triple entendre on the classic joke.

% \begin{Verbatim}
facebook wall posts, 4-grams
100 top paths:
0.002525504 pittsburgh
0.002093510 steelers
0.002093510 sfo
0.001953943 it's
0.001256106 bdl
0.001116539 sigbovik
0.001099196 facebook
0.000976971 mic
0.000976971 s
0.000976971 can't
0.000837404 i'm
0.000837404 icfp
0.000837404 app
0.000697837 x
0.000697837 drunj
0.000697837 g
0.000616021 ther
0.000558269 doesn't
0.000558269 's
0.000558269 sr
0.000558269 pm
0.000558269 fyi
0.000558269 wtf
0.000558269 k
0.000558269 don't
0.000558269 today's
0.000558269 m
0.000558269 what's
0.000456766 iphone
0.000429438 inted
0.000418702 oh
0.000418702 phl
0.000418702 gonna
0.000418702 pre
0.000418702 zrh
0.000418702 th
0.000418702 h
0.000418702 b
0.000418702 n
0.000418702 ok
0.000418702 d
0.000418702 that's
0.000418702 rex
0.000418702 fanzibar
0.000418702 st
0.000418702 kinda
0.000418702 didn't
0.000355262 cally
0.000334962 brillobox
0.000314027 superburrito
0.000305304 weath
0.000279135 techno
0.000279135 vii

>>>>

% \usepackage{eclbkbox}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{code}
% \usepackage{amscd}
% \usepackage{xy}
\usepackage{graphicx}
% \usepackage{fancyhdr}
% \usepackage{color}
% \usepackage[dark,all,bottom,landscape,timestamp]{draftcopy}
% \usepackage{everypage}

\pagestyle{empty}

@@

\date{1 April 2011}

@@

\begin{abstract}
This paper is an earnest attempt to answer the following question
scientifically: What words ought to exist?
\end{abstract}

\vspace{1em}
{\noindent \small {\bf Keywords}:
 computational cryptolexicography, n-Markov models, coinduction
}

\section*{Introduction}
During a recent high-stakes game of Scrabble-brand Crossword
Puzzle\footnote{Scrabble is a registered trademark of Hasbro
  Inc./Milton Bradley, and Mattel/JW Spear \& Sons plc.} I had what
could only be described as a killer bingo word (all 7 tiles) that,
after careful study, I determined could not be placed anywhere on the
board. Later in that same game, I had another sequence of letters that
just totally seemed like it should be able to make some long-ass
words, like for example ``oilsoap'' which turns out is not a legal
Scrabble word.\!\footnote{There are actually no 7-letter words that
  can be made from these letters. Don't even bother. Even if playing
  off an existing letter on the board, the best we can do are the
  non-bingos ``topsoil,'' ``topsail,'' or ``poloist'' with an
  available {\it t}.} This naturally made me frustrated and I wanted
to do something about it. Why can't ``oilsoap'' be a word? Or
``loopsia''? Words are introduced into the lexicon all the time. My
first reaction of course was to make an online version of Scrabble
where all words are legal. This is called Scrallbe (where they can
{\it all be} words!\footnote{As of 2011, the official Scrabble slogan
  is ``every word's a winner!'' which is clearly false.}) This is
available at {\tt http://snoot.org/toys/scrallbe}, and is pretty
boring, I gotta be honest (Figure~\ref{fig:scrallbe}).

\begin{figure}
\begin{center}
\includegraphics[width=0.75 \linewidth]{scrallbe-screenshot}
\end{center}\vspace{-0.1in}
\caption{In-progress Scrallbe game, 753 points.}
\label{fig:scrallbe}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\linewidth]{wishlist-cdf}
\vspace{-0.1in}\caption{Cumulative distribution of word frequency.
  Approximately 25,000 different words (y axis) were issued 55 times
  or fewer (x axis). The ``total'' area does not appear much larger
  than its components because this is a log-log plot.}
\label{fig:gamedistribution}
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\linewidth]{coin1}
\caption{Markov chain for the SOWPODS word list, where darker squares
indicate higher probability. The darkest is the transition from {\it q}
to {\it u} ($98\%$), which is not surprising.}
\label{fig:sowpodsbigrams}
\end{figure}

The next step is to build a Markov chain for a list of real words and
see what results. I built one for the SOWPODS word list, which results
in the table in Figure~\ref{fig:sowpodsbigrams}. These
are the most likely words, with real words filtered out:

% XXX improve layout
\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
4.99\%  & s     &   0.17\%  & y   \\
1.75\%  & d     &   0.17\%  & p   \\
0.95\%  & g     &   0.16\%  & a   \\
0.55\%  & c     &   0.16\%  & n   \\
0.43\%  & r     &   0.15\%  & ps  \\
0.42\%  & t     &   0.13\%  & ms  \\
0.40\%  & e     &   0.13\%  & ts  \\
0.35\%  & m     &   0.13\%  & ds  \\
0.32\%  & ss    &   0.11\%  & hy  \\
0.20\%  & rs    &   0.11\%  & k   \\
0.19\%  & h     &   0.11\%  & ng  \\
0.18\%  & l     &   0.11\%  & ly  \\
\end{tabular}
\end{center}

@@

There is a standard solution to this problem, which is to generalize
the Markov chain to keep more than one letter of history. So instead
of just tallying how often {\it g} follows {\it n}, we count how often
{\it g} follows {\it in} (and any other pair of letters).\!\footnote{
  The details are straightforward, except possibly that we now imagine
  each word to start with two (or in general, $n$) copies of the start
  symbol, so that we see ``\<\<helpful\>''. The column corresponding
  to the history {\bf \<\<} tells us the frequency of letters that
  start words, and for example the column {\bf \<h} tells us the
  frequency of letters that follow {\it h} when it appears at the
  start of a word. We do not need to repeat the ending character \>
  because once we see it, we never do anything but end the word.} This
makes the table pretty large, so you'll just have to look at
Figure~\ref{fig:sowpodsbigrams} again and imagine it being 28 times
wider. But the good news is that it invents much better words:

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{Markov chain with $n=2$.} \\
\hline
.709\% & ing     &   .110\% & le    \\
.248\% & ses     &   .107\% & der   \\
.169\% & des     &   .107\% & ove   \\
.154\% & nes     &   .101\% & gly   \\
.140\% & sts     &   .088\% & hy    \\
.131\% & se      &   .085\% & ung   \\
.128\% & ings    &   .083\% & cy    \\
.126\% & ded     &   .081\% & pres  \\
.117\% & cal     &   .080\% & pers  \\
\end{tabular}
\end{center}

These are even, like, pronounceable. The best news is that they keep
getting better the more history we keep:

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{Markov chain with $n=3$.} \\
\hline
.109\% & des     &   .038\%  & ent     \\
.078\% & pers    &   .036\%  & dist    \\
.076\% & cal     &   .035\%  & ble     \\
.062\% & pres    &   .035\%  & ches    \\
.045\% & nons    &   .034\%  & gly     \\
.044\% & ress    &   .034\%  & inted   \\
.042\% & ing     &   .034\%  & dists   \\
.040\% & pred    &   .033\%  & lity    \\
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{Markov chain with $n=4$.} \\
\hline
.045\% & unders  &   .017\% & heters  \\
.034\% & dising  &   .016\% & sters   \\
.029\% & pers    &   .015\% & stic    \\
.028\% & cally   &   .014\% & pering  \\
.023\% & inted   &   .013\% & dises   \\
.020\% & heter   &   .013\% & ching   \\
.019\% & tric    &   .012\% & shing   \\
.018\% & ster    &   .012\% & dest    \\
.018\% & hier    &   .011\% & teless  \\
.018\% & unded   &   .011\% & resis   \\
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{Markov chain with $n=5$.} \\
\hline
\multicolumn{4}{l}{{\tt GetTempFileName failed with error 5}} \\
\end{tabular}
\end{center}

With four letters of history, the words produced are quite good! (The
results at $n=5$ are somewhat disappointing since the program crashes
from running out of memory. The table at $n=5$ would have over 481
million entries.) Many of these seem like real words. Some even
suggest meaning because they contain common morphemes. To make the
case that these are not just real-looking words but characteristic of
the English language, compare the results of the same algorithm on the
dictionary from the Italian language edition of Scrabble, which is
probably called {\it Scrabblizzimo!} (Figure~\ref{fig:italian}).
Italian is lexicographically a more compact language than English
(Figure~\ref{fig:italianbigrams}); there are only 21 letters (outside
of occasional interlopers in loan words like {\it jeans} and {\it
  taxi}). Moreover, even though the dictionary contains 585,000 words
(twice as many as English), the probabilities of observing these
non-words are much higher than the most likely English ones.

\begin{figure}[t]
\includegraphics[width=\linewidth]{italiancoin1}
\caption{Markov chain for the Italian language. Again darker cells
  indicate higher probability. Italian has more lexicographic
  structure recognizable from bigraphs than English does: Note that
  the extremely rare letters ``j'', ``k'', ``q'', ``w'', ``x'', and
  ``y'' have almost empty rows. ``z'' very frequently follows ``z'',
  as in {\it pizza}. Words almost always end in a vowel.}
\label{fig:italianbigrams}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
.137\%  &  ammo       &   .026\%  &  rino      \\
.071\%  &  rice       &   .025\%  &  diste     \\
.061\%  &  rico       &   .024\%  &  risti     \\
.055\%  &  este       &   .023\%  &  disci     \\
.053\%  &  scono      &   .022\%  &  riasse    \\
.049\%  &  immo       &   .022\%  &  riassi    \\
.047\%  &  assero     &   .021\%  &  cate      \\
.047\%  &  scano      &   .019\%  &  rite      \\
.038\%  &  rammo      &   .019\%  &  cando     \\
.034\%  &  cata       &   .018\%  &  riassero  \\
.034\%  &  assimo     &   .018\%  &  riassimo  \\
.032\%  &  riate      &   .018\%  &  dete      \\
.032\%  &  disce      &   .018\%  &  disca     \\
.030\%  &  esti       &   .017\%  &  risca     \\
.029\%  &  rica       &   .017\%  &  cente     \\
.028\%  &  endo       &   .016\%  &  acci      \\
.027\%  &  dissimo    &   .015\%  &  centi     \\
.026\%  &  rici       &   .015\%  &  girono    \\
\end{tabular}
\end{center}
\caption{Most probable words induced by the Markov Markov chain for
  the Italian language ($n=4$).}
\label{fig:italian}
\end{figure}

\subsection{Usage-weighted methods}

One criticism of this approach is that it considers every word in the
word list to be equally important.\!\footnote{In fact, the common part
  of words with many different conjugations is in essence counted many
  times. This means {\it ornithology} in its six different forms
  contributes six times as much to our model as the word {\it the}!} I
object on the philosophical grounds that some words that already exist
{\em ought to exist more} than other words that already exist. For
example, {\it congenital} is a much nicer word than the plain ugly
{\it congenial}, and is reflected by the fact that {\it congenital} is
used five times more frequently than {\it
  congenial}.\!\footnote{14,200,000 times to 2,820,000, on the
  Internet, according to Google.} In this section, we produce Markov
models of words weighted by the frequency with which people tend to
use them. This is just a simple matter of training the model on some
natural language corpus (with many occurrences of each word, or no
occurrences of unpopular words) rather than a flat list of all alleged
words.

{\bf Facebook.}\ Since the best research is intensely navel-gazing, I started by 
analyzing a corpus of my own writing, specifically my Facebook status
updates since March~2006. There were 1,386 status updates containing
such jems as ``Tom Murphy VII thinks mathfrak is straight ballin''
and ``Tom Murphy VII global L\_50 reused for unused\_36166!!''. The
most likely words with $n=4$:

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{My Facebook status updates, $n=4$.} \\
\hline
.252\% &  pittsburgh   & .097\% &  can't    \\   
.209\% &  steelers     & .083\% &  i'm      \\   
.209\% &  sfo          & .083\% &  icfp     \\   
.195\% &  it's         & .083\% &  app      \\   
.125\% &  bdl          & .069\% &  x        \\   
.111\% &  sigbovik     & .069\% &  drunj    \\   
.109\% &  facebook     & .069\% &  g        \\   
.097\% &  mic          & .061\% &  ther     \\   
.097\% &  s            & .055\% &  doesn't  \\
\end{tabular}
\end{center}

This is the worst. Not only does it contain loads of one-letter words that
we have already determined are verboten,\!\footnote{Note that since $n=4$ these
words have to actually appear in status updates to have nonzero probability for
this list. ``g'' is explained by frequent occurrences of ``e.g.'', for example.}
but the rest are just non-words that I tend to use like the names of cities,
prestigious conferences, or IATA airport codes. The main problem is that there
is simply not enough data from which to generalize.

{\bf Wikipedia.}\ I tried again, but with Wikipedia, using a snapshot
of the English site from June~2009. This is 23 gigabytes of data, most
of it expository text composed by native speakers, plus bathroom humor
vandalism. The list produced by this analysis is much better, though it
contains artifacts from non-English Wiki language used in articles. The
unabridged list appears in the appendix; my hand-selected favorites:

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{English Wikipedia, $n=3$.}     \\
\hline
.0287\% & smally      &     .00518\% & reporth    \\
.0156\% & websity     &     .00484\% & delection  \\
.0156\% & stude       &     .00459\% & grounty    \\
.0124\% & chool       &     .00437\% & betweek    \\
.0120\% & fontry      &     .00431\% & fination   \\
.0102\% & undex       &     .00388\% & manuary    \\
.0099\% & octory      &     .00360\% & whicle     \\
.0096\% & coibot      &     .00262\% & stategory  \\
.0084\% & footnot     &                           \\
\end{tabular}
\end{center}

Lots of these could be the names of tech startups or Pok\'emon.

% laplace smoothing?

@@

In the earlier sections I blithely produced tables of the most
probable words according to an $n$-Markov chain. It is not obvious how
to do this (or that it is even possible), so I explain the algorithm
in this section. It's safely skippable, I mean if you don't want to
know about a pretty cool algorithm that's not that complicated and
might even be new, plus {\em dual math}.

Computing the probability of an individual word is easy. We prefix it
with $n$ copies of the start symbol \<, suffix it with a single \>,
and then look up the probability of each symbol given its $n$
preceding symbols in the table, and multiply those all together. We
can compute the probability of any word this way. The problem with
sorting all of the possible words by their probabilities is that there
are an infinite number of them. We can't just look at short words
first, either, because for example the word ``thethethe'' is many
times more likely ($p = 6.08\times 10^{-11}$) than the shorter
``qatzs'' ($9.07\times 10^{-12}$).

The solution is to use coinduction. Most people remember induction
from school, maybe, which is the one where you have some base case
like ``0 is even'', and then you prove that all numbers are either
even or odd by assuming ``$n - 1$ is even or odd'' and proving ``$n$
is even or odd''. From this we conclude that every number is either
even or odd. The idea is the proof shows how to, for any given number
$m$, count down to the base case ``0 is even'', and then repeatedly
apply the $n-1$ step (inductive step) to get back up to $m$. This is a
great way to prove facts about finite things like numbers. Think of
induction as a way you prove a statement like ``Good to the last
drop,'' or ``There's always room for Jello.''

Coinduction is a good proof technique for infinite things, like a
sorted infinite list of possible strings. The idea behind coinduction
is kind of like, you prove something like ``0 is a number'' (the base
case), then prove something like ``if $n$ is a number, then $n+1$ is a
larger number'', and then conclude that there exists an infinite
series of numbers, each larger than the previous one. Think of
coinduction as a way you prove a statement like ``Once you pop, you
can't stop,'' or ``Never gonna give you up.''

To sort the infinite list we don't actually use coinduction (we're
not going to prove anything, just implement it), but its computational
counterpart, corecursion. I just can't resist the ``coin'' pun.

What we do is define a function ``most probable paths'', which returns
a (possibly infinite) stream of strings for a given starting state.
Each string is finite and ends with the terminal symbol \>, and they
appear sorted by decreasing probability. (The most probable words
overall will be just the first elements from the stream returned by
this function when using a starting state like \<\<\< for $n=3$.)
Since we don't want to explore all possible strings in order to
produce this list (there are infinitely many), the trick is to put a
lower bound on the probability of the words that will be included.
There are always finitely many words with probability greater than a
given positive value, unless the Markov chain contains a cycle where
each edge has probability 1. (This is impossible for Markov chains
created only by observing finite strings, such as all the ones in this
paper.) It is efficient to use a very small lower bound with this
algorithm, like 0.00000000000001.

So the specification for ``most probable paths'' is to return all of
the strings (that end with \>) that exceed the given lower bound in
probability, sorted in descending probability order. It is easy to
check the path directly to \>; we compare its probability to the lower
bound by just looking it up in the table, and consider it if it
exceeds the lower bound. For any other symbol {\it sym}, we will
proceed (co)recursively: Call the probability of seeing {\it sym} next
{\it p}, and then compute {\it tails}, all of the most probable paths
starting in the state we would be in upon seeing {\it sym}. We turn
{\it tails} into the sorted stream for the current state by just
adding {\it sym} to the beginning of each string in it, and
multiplying the probability by {\it p}. It remains sorted because
multiplying by the same {\it p} is monotonic. The most important
thing, which makes the algorithm practical (indeed terminate at all),
is that we pass in a new lower bound: The current lower bound divided
by {\it p}. After all, the outputs will be multipled by {\it p}, so
they have to exceed this in order to meet the lower bound. This tends
to increase the lower bound (sometimes over 1) since probabilities are
between 0 and 1. This way, we only need to search a few symbols deep
before it's clear that no string can exceed the lower bound.

Now we have a list of sorted streams, at most one for each symbol in
our alphabet. It is fairly straightforward to merge these into a
single sorted stream, by only looking at the first element from each
one. Pseudocode for {\tt most\_probable\_paths} appears in
Figure~\ref{fig:mostprobablepaths} and for {\tt merge\_sorted} in
Figure~\ref{fig:mergesorted}. Performance of this code is great;
building the Markov chains (or even just reading the dictionary files)
dominates the latency of the analyses in this paper.

\begin{figure*}
\begin{code}
fun most\_probable\_paths \{ lower_bound : real, state : state \}
       : \{ string : symbol list, p : real \} stream =
  let
    fun nexts i =
      case symbol\_from\_int i of 
      NONE => nil
    | SOME sym =>
      let 
        val p = (* probability of seeing sym in this state *)
      in
        if p < lower\_bound
        then nexts (i + 1)
        else if sym = end\_symbol
             then S.singleton \{ string = nil, p = p \} :: nexts (i + 1)
             else 
               let
                   val lb' = lower\_bound / p
                   val tails =
                       most\_probable\_paths \{ lower\_bound = lb',
                                             state = advance\_state (state, sym) \}
               in
                   (* Now multiply through the probabilities and add the symbol
                      to the head of the strings. *)
                   Stream.map (fn \{ string = t, p = p' \} =>
                               \{ string = sym :: t, p = p * p' \}) tails ::
                   nexts (i + 1)
               end
      end

    (* Try all next symbols. *)
    val streams = nexts 0
  in
    S.merge\_sorted bysecond\_real\_descending streams
  end
\end{code}
\caption{Pseudocode for {\tt most\_probable\_paths}. {\tt advance\_state}
gives a new state from a previous state and symbol observed, so that
for example {\tt advance\_state(abc, z)} gives {\tt bcz}. The pseudocode
for {\tt merge\_sorted} is given in Figure~\ref{fig:mergesorted}.}
\label{fig:mostprobablepaths}
\end{figure*}

\begin{figure*}[t]
\begin{code}
fun merge\_sorted cmp l =
  let
      fun ms nil () = Nil
        | ms (s :: t) () =
          case force s of
              Nil => ms t ()
            | Cons (v, ss) =>
              ms\_insert v [ss] t

      and ms\_insert bv sg nil =
          Cons (bv, delay (ms sg))
        | ms\_insert bv sg (s :: t) =
          case force s of
              Nil => ms\_insert bv sg t
            | Cons (v, ss) =>
                case cmp (bv, v) of
                    GREATER =>
                      ms\_insert v (singleton bv :: ss :: sg) t
                  | \_ => ms\_insert bv (s :: sg) t
  in
      delay (ms l)
  end
\end{code}
\caption{Pseudocode for {\tt merge\_sorted}. {\tt ms} merges a
sorted list, and {\tt ms\_insert} is a helper where we have a
candidate best value {\tt bv} which will either be the one we
return at the head of the stream, or we'll replace it and then
stick {\tt bv} somewhere to be returned later. (This algorithm
can be improved by making a data structure like a (co)heap;
this is just a simple first pass.)}
\label{fig:mergesorted}
\end{figure*}

\section{Special cases}

The empty string?? Is that a word? Could it be? Dude that is blowing my mind.

\section{Backformation} \label{sec:backformation}

The lexicon is generative, in the sense that it's possible to make new
words that are generally acceptable, by following rules. Most people
recognize pluralization of nouns by adding {\it --s} (even for novel
words), or adding prefixes like {\it anti--}. We could investigate
words that ought to exist by the application of rules, such as {\it
  examplelikelikelikelikelikelike}, but I see no straightforward way
to justify the relative strength of such words.

A related way for words to enter the lexicon is by backformation. This
is the reverse of the above process: A word like {\it laser}
(initially an initialism) is legal, and then by running the rules of
English backwards, we start to use {\it lase} as a word (the verb
that a laser most frequently applies). In this section, I attempt
to determine formation rules in English (by simple lexical analysis
of the set of legal words) and then run these rules backwards to find
words that seemingly should already exist.

{\bf Prefixes and suffixes.}\ The first order of business is to find
prefixes and suffixes that are usually modular. The kind of thing
we're tring to find are ``anti--'' and ``--ing''; stuff you can often
add to a word to make a related word. The approach is straightforward.
For each word, consider splitting it at each position. For {\it dealing},
we have {\it d/ealing}, {\it de/aling}, etc. For every such split,
take the prefix (e.g. ``de'') and remainder (``aling''); if the remainder
is still a legal word, then the prefix gets one point. {\it aling} is not
a word so no points here for ``de''. We also do the same thing for suffixes
(using the exact same splits, symmetrically). In this case we'll only get
points for ``--ing'' since {\it deal} is a word. Every time a prefix or
suffix appears we test to see if it is being applied modularly, and
the final score is just the fraction of such times. Here are the ones with
the highest scores:

\begin{verbatim}
1.000000000 -zzyingly    1/1
1.000000000 -zzying      1/1
1.000000000 -zzuolanas   1/1
1.000000000 -zzuolana    1/1
1.000000000 -zzotints    1/1
1.000000000 -zzotintos   1/1
1.000000000 -zzotinto    1/1
1.000000000 -zzotinting  1/1
...
\end{verbatim}

Well, it's good to know that 100\% of the time, you can remove
``--zzotinting'' from a word and it will still be a word. But this
inference is supported by just one observation (the word {\it
  mezzotinting}); there are actually hundreds of such unique prefixes
and suffixes. We need a better list.\!\footnote{The right thing to do
  here is probably to use binomial likelihood rather than the
  scale-independent fraction. But simpler approaches produce pretty
  good lists.} Removing the ones that appear just a single time
doesn't really help that much:

\begin{verbatim}
1.000000000 -zzazzes     3/3
1.000000000 -zzazz       3/3
1.000000000 -zzans       3/3
1.000000000 -zzanim      2/2
1.000000000 -zzan        3/3
1.000000000 -zygotic     3/3
\end{verbatim}

Still bad. Let's turn up the juice to prefixes and suffixes that
appear at least 10 times.

\begin{verbatim}
1.000000000 -wrought    10/10
1.000000000 -writings   12/12
1.000000000 -wraps      10/10
1.000000000 -wrap       11/11
1.000000000 -worms      69/69
1.000000000 -worm       69/69
1.000000000 -working    21/21
\end{verbatim}

Much better! But the next step is going to be to try removing these
prefixes and suffixes from words that have them, to find new words.
Since these have modularity of 100\%, we already know that every
time we apply them, the result will already be a word. So they are
useless for our analysis. Here are the most modular prefixes
and suffixes with modularity {\it strictly less than} 1.

\begin{verbatim}
0.985714286 -makers     69/70
0.985714286 -maker      69/70
0.983606557 -wood       120/122
0.983471074 -woods      119/121
0.982758621 -down       57/58
0.982658960 -works      170/173
0.981818182 -houses     108/110
0.981818182 -house      108/110
0.981132075 kilo-       52/53
0.980752406 -less       1121/1143
0.980743395 over-       2190/2233
0.980000000 -books      49/50
0.980000000 -book       49/50
0.979591837 -proof      48/49
0.979310345 -lessnesses 142/145
0.979069767 -ships      421/430
0.978723404 -lessness   184/188
0.978723404 -board      138/141
0.978494624 -woman      91/93
0.978021978 -women      89/91
0.977528090 -ship       435/445
0.977272727 -manship    43/44
0.976744186 -weeds      84/86
0.976470588 after-      83/85
0.976190476 -manships   41/42
0.976190476 -making     41/42
0.976190476 -craft      41/42
0.976190476 -boats      41/42
0.976190476 -boat       41/42
\end{verbatim}

Wow, now we're talking! The single word that cannot have ``--maker''
removed is {\it comaker}, suggesting that {\it co} should be word
(noun: ``What a comaker makes.'').

Given this list, the next step is to identify potential words that
can be backformed by removing prefixes or adding suffixes from existing
words. Such a string can often be found via multiple prefixes and
suffixes. For example, {\it twing} can be formed by removing ``--ing''
from {\it twinging} (a false positive, since the root word is actually
{\it twinge} in this case) as well as by removing the prefix ``lef--'',
which has modularity of 20\% (including splits such as ``lef/tie'').
Maybe not good justification, but {\it twing} is a pretty good word
anyway.

We define the probability of a word as its Markov probability (with
$n=4$, as this seems to produce the best results), times the
probability that at least one of the potential backformation rules
applies.\!\footnote{As above we only allow backformation rules that
  have at least 10 occurrences, to prevent degeneracy.} Here are the
most likely words by backformation:

\begin{verbatim}
word    prob   most likely backformation rules
==============================================
dises   .023%  para- (0.42) fluori- (0.39) 
               melo- (0.35) bran- (0.31)  
tring   .020%  hams- (0.36) scep- (0.35) 
               bows- (0.33) hearts- (0.29)  
disms   .017%  triba- (0.31) drui- (0.30) 
               bar- (0.27) invali- (0.27)  
ching   .017%  day- (0.86) hot- (0.69) 
               star- (0.51) guillo- (0.50)    
sking   .017%  dama- (0.24) imbo- (0.18) 
               fri- (0.18) atta- (0.17)     
cally   .015%  anti- (0.78) specifi- (0.61)
               magnifi- (0.55) phoni-    
pring   .015%  days- (0.67) heads- (0.62)
               outs- (0.54) ups- (0.51)    
\end{verbatim}

I think that this approach shows promise, but there appear to be a few
problems: Many of these ``rules'' can be explained by bad segmentation
(``heads--'' appearing to be modular, for example, is really just
``head--'' plus ``s'' being a common letter.) Second, I believe the
disjunctive probability of any rule applying is too naive for
determining the score. For example, {\it tions} has almost a thousand
different prefixes that could apply to it; the chance of {\it any one}
of them applying is very nearly 1. But this is actually because ``tions''
is just a common way for a word to end. Legitimate root words to which
many good prefixes are applied cannot be easily distinguished from
common suffixes by this symmetric algorithm. More work is called for
here.

@@

On occasion I have been accused of ``overthinking'' problems, whatever
that means. So to compare, I next hazarded a tried and true technique
from grade school, the survey.

I asked a few people who happened to be around, ``What word ought to
exist?'' Most people did not know what to make of this question, and
also, because people seem to revel in the opportunity to get (well
deserved) revenge on me by being disruptive trolls, many of the
answers were designed to be unusable. In order to not reprint
everyone's bullshit---but not introduce bias by selectively removing
data---I discarded random subsets of the data until it did not contain
bullshit any more.

\begin{center}
\begin{tabular}{rl}
Rob: &  etsy, nuuog \\
Chris: &  nurm \\
David: &  wafflucinations \\
Lea: &  hnfff \\
Reed: &  pansepticon \\
Jessica: &  gruntle \\
\end{tabular}
\end{center}

From this we can conclude that 16\% of people wish {\it nurm} were a word, and so on.
These words did not come with definitions, except for {\it gruntle}, which Jessica
gives as ``the opposite of disgruntle''. This is actually already a word, but
it was the inspiration for Section~\ref{sec:backformation}. {\it etsy} is the name
of a popular on-line crafts community so I don't know why Rob would suggest that.
The meaning of {\it wafflucinations} is clear from morphological analysis.

% TODO: length normalization. just take the average (or max?)
% probability for a word of length n, and normalize by that (which means what?)?

@@

\section{Conclusion}

In this paper I investigated several different ways of answering the
question: What words ought to exist? Each method produces different
words, and some don't work that well, but nonetheless we have several
rich sources of words, each time scientifically justified. I conclude
with a section of recommendations for words that ought to exist, along
with definitions.

\subsection{Recommendations}

{\it Sweeeeeeeeeeeeeeeeeeet} with 19 {\it e}s is the clear favorite
based on analysis of usage, so this one should be introduced. It means
``Really sweet.''

{\it Rane} sounds too much like {\it rain}, but {\it sare} has a
unique pronunciation and many people seem to think it's already a
word. I propose that {\it sare} be introduced as a noun meaning, ``a
word that sounds real but isn't.''

{\it Cho} is similarly easy to pronounce and spell. I propose that it
be defined as ``A kind of cheese,'' so that we can really nail the new
triple entendre on the classic joke. {\it Chomaker} is someone who
makes that kind of cheese.

{\it Unders} was one of the most frequently occuring words towards the
top of many analyses. This word should be a colloquialism for
underwear, which would probably already be understood from context.

{\it Dise} is suggested by both the Markov model (as {\it dises}, {\it
  dising}) and backformation (as {\it dises}). I like thinking of it
as being the root of {\it paradise}, where {\it para--} means
something like ``along side of'' or ``resembling''. So {\it dise} is
the place you're really looking for when you get to paradise and
realize it's just a mediocre country club.

{\it Helafelo} is one hell of a fellow.

\medskip
{\bf Addendum.} During the preparation of this paper, the Scrallbe
game has converged on a culture where the words played are real-seeming,
with creative definitions. Examples: {\it frodeo} (``Gandalf is the
clown.'') {\it pridefax} (``An unproven treatment for telephone
anxiety.'') {\it eeeee} (``eeeee'') {\it orzigato} (``Move mr. roboto.
for great justice.'') {\it stovebed} (``Now you don't have to get out
from under the covers to make breakfast.'') {\it ovawiki} (``The free
egg cell that anyone can edit.'') {\it gaptave} (``Two discontinuous
musical intervals.'') {\it achoolane} (``Nostril (colloq.)'') {\it
  gplerious} (``Completely overcome by software licenses.) {\it
  bestcano} (``Some eruptions are better than others.'') Thanks to the
players for their contributions, especially Ben~Blum, Chrisamaphone,
Rob~Simmons, and Flammin~Fingers.
